---
title: "ESM 244 Lab 1 Satie"
author: "Satie Airame"
date: "1/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(sf)
library(tmap)
```

### Read in SF trees data

```{r}
sf_trees <- read_csv(here("data", "sf_trees", "sf_trees.csv"))
```

### Basic wrangling reminders

Refresh some skills for data wrangling and summary statistics using functions in the 'dplyr' package.
Find the top five highest observations of trees by legal_status, do some wrangling, make a graph.

```{r}
top_5_status <- sf_trees %>% 
  count(legal_status) %>%  # recognize groups, finding counts, in a table
  drop_na(legal_status) %>%  # to remove missing values labeled NA
  rename(tree_count = n) %>% # new name goes on the left, old name on the right
  relocate(tree_count) %>%  # moves tree_count to first column position
  slice_max(tree_count, n = 5) # indicate in which colunn to look and the number you want to present, there is also slice_min available
```
### Make a graph of top five observations by legal status

```{r}
ggplot(data = top_5_status, aes(x = fct_reorder(legal_status, tree_count), y = tree_count)) + # uce fct_reorder to reorganize data in order of tree_count, alternative to use convert to factor and use fct_reorder
  geom_col() +
  labs(x = "Legal Status", y = "Tree Count") +
  coord_flip() + # flip the axes in case the names are too long on x axis
  theme_minimal()

```

### A few more data wrangling refresher examples
Only keep observations (rows) for Blackwood Acacia Trees

```{r}
blackwood_acacia <- sf_trees %>% 
  filter(str_detect(species, "Blackwood Acacia")) %>% # looks within a variable (sf_trees data) that we specify to identify if there are any strings there, does not need to be a complete match, just part of the string
  select(legal_status, date, latitude, longitude) # pick columns that you want to pick or exclude

ggplot(data = blackwood_acacia, aes(x = longitude, y = latitude)) +
  geom_point()
```

### tidyr::separate() and unite() functions
Useful for combining and separating columns

```{r}
sf_trees_sep <- sf_trees %>% 
  separate(species, into =
c("spp_scientific", "spp_common"), sep = "::") # separate one column into two columns using data within that column, create a vector c with two new columns, then name the separator (::)
```

Exampe: tidyr::unite()


```{r}
sf_trees_unite <- sf_trees %>% 
  unite("id_status", tree_id:legal_status, sep = "_cool!_") # combine two columns into one and add a separator ("_cool!_"), could use a vector or use a colon, which indicates "from this column to that column"
```













